---
title: "Practical Machine Learning Course Project"
author: "kuna Abhinav 190909072"
date: "2022-10-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Loading Data and Libraries
Loading all the libraries and the data
```{r results='hide', message=FALSE}
library(lattice)
library(ggplot2)
library(caret)
library(kernlab)
library(rattle)
library(corrplot)
set.seed(1234)
```

```{r}
traincsv <- read.csv("./data/pml-training.csv")
testcsv <- read.csv("./data/pml-testing.csv")
dim(traincsv)
dim(testcsv)
```

We see that there are 160 variables and 19622 observations in the training set, while 20 for the test set.


# Cleaning the Data

Removing unnecessary variables. Starting with N/A variables.
```{r}
traincsv <- traincsv[,colMeans(is.na(traincsv)) < .9] #removing mostly na columns
traincsv <- traincsv[,-c(1:7)] #removing metadata which is irrelevant to the outcome
```

Removing near zero variance variables.
```{r nzv}
nvz <- nearZeroVar(traincsv)
traincsv <- traincsv[,-nvz]
dim(traincsv)
```

Now that we have finished removing the unnecessary variables, we can now split the training set into a **validation** and sub **training** set. The testing set "testcsv" will be left alone, and used for the final quiz test cases. 
```{r}
inTrain <- createDataPartition(y=traincsv$classe, p=0.7, list=F)
train <- traincsv[inTrain,]
valid <- traincsv[-inTrain,]
```


# Creating and Testing the Models
Here we will test a few popular models including: **Decision Trees**, **Random Forest**, **Gradient Boosted Trees**, and **SVM**. This is probably more than we will need to test, but just out of curiosity and good practice we will run them for comparison.

Set up control for training to use 3-fold cross validation. 
```{r}
control <- trainControl(method="cv", number=3, verboseIter=F)
```

## Decision Tree

**Model:** 

```{r, cache=TRUE}
mod_trees <- train(classe~., data=train, method="rpart", trControl = control, tuneLength = 5)
fancyRpartPlot(mod_trees$finalModel)
```

**Prediction:**

```{r}
pred_trees <- predict(mod_trees, valid)
cmtrees <- confusionMatrix(pred_trees, factor(valid$classe))
cmtrees
```

## Random Forest

```{r, cache=TRUE}
mod_rf <- train(classe~., data=train, method="rf", trControl = control, tuneLength = 5)
pred_rf <- predict(mod_rf, valid)
cmrf <- confusionMatrix(pred_rf, factor(valid$classe))
cmrf
```

## Gradient Boosted Trees

```{r, cache=TRUE}
mod_gbm <- train(classe~., data=train, method="gbm", trControl = control, tuneLength = 5, verbose = F)
pred_gbm <- predict(mod_gbm, valid)
cmgbm <- confusionMatrix(pred_gbm, factor(valid$classe))
cmgbm
```


# Predictions on Test Set

Running our test set to predict the classe (5 levels) outcome for 20 cases with the **Random Forest** model.
```{r}
pred <- predict(mod_rf, testcsv)
print(pred)
```

#Correlation matrix of variables in training set
```{r}
corrPlot <- cor(train[, -length(names(train))])
corrplot(corrPlot, method="color")
```

Plotting the models
```{r}
plot(mod_trees)
plot(mod_rf)
plot(mod_gbm)
```

